### 2014-03-13 src/methodology.Rnw
udpate

\section{Methodology}
\label{sec:methodology}
Our selection of programs was driven by a few overriding concerns. Our primary
requirement was that, our results had to be as widely applicable as possible
for real-world programs. Secondly, we strived for a statistically significant
result, reducing the amount of variables present in the experiments. For this
reason we chose a random sample of Java projects from Github\cite{github}
that uses the popular maven\cite{maven} build system, that has been used
elsewhere\cite{gopinath2014code}. We utilize two flavors of
of test-suites. The original test-suites present in a limited number of
projects, and automatically generatd test-suites using randoop.  Due
limited resources, we had to further reduce the programs and test-suites
to those that could finish the mutation run in 10 minutes. From this set
of test-suites, we sampled 110 suites each for both original and randoop
generated test-suiets. This gave us 81 unique projects for both original
and randoop, and remaining projects having both flavors of test suites.

@@

For a detailed description of each mutants, please see~\cite{pitmut}.
PIT was also modifed to provide random selection of mutants according
to various criterias. Finally, to remove random noice, each criteria
described has been run four times, and the results are averaged to produce
the final result.

\begin{table}
\small{
\begin{tabular}{ll}
\hline \hline \\
\pitin & Remove negative sign from numbers\\
\pitrv & Mutate return values\\
\pitm & Mutate arithmetic operators\\
\pitvmc & Remove void method calls\\
\pitnc & Negate conditional statements\\
\pitcb & Modify boundaries in logical conditions\\
\piti & Modify increment and decrement statements\\
\pitnvmc & Remove non-void method calls, returning default value \\
\pitcc & Replace constructor calls, returning null \\
\pitic & Replace inline constants with default value \\
\pitri & Remove increment and decrement statements \\
\pitemv & Replace member variable assignments with default value \\
\pites & Modify switch statements \\
\pitrs & Replace switch labels with default (thus removing them) \\
\pitrc & Replace boolean conditions with true\\
\pitnc & Replace boolean conditions with false\\
\hline
\end{tabular}
}
\caption{PIT Mutation Opeators (We have used abbreviations instead of operator names.)}
\label{tbl:pitop}
\end{table}

One of the criterias we sampled was \textit{statement deletion}.
Since we use PIT, which is a bytecode based mutation engine, implementing direct equivalent of
source code statement deletion is very hard. We were able to approximate statement deletion
by combining different bytecode operators, which is explained in Subsection~\ref{subsec:stmtdel}.

\subsection{Sampling Criterias}

@@

Next, we explored in the other direction, by forcing the program elements
to return atleast one element by using the \textit{ceil(sample)}. That is,
irrespective of the fraction being sampled, the simplest elements always
contributed atleast one mutant to the total sample.

@@

\subsubsection{Statement Deletion:}
\label{subsec:stmtdel}
The basic statement deletion was modeled on the work by Deng et
al.~\cite{deng2013empirical}. The operations on single statements was modeled
using \pitvmc, \pitnvmc, \pitcc, \pitemv, \pitri
for simple statements, and using \pitrc for control structures,
which replaced the boolean condition with false, resulting in removal of
the block of statements. The operator for return values was modeled using
\pitrv, which is similar.  The operators for \textit{while,for},
and \textit{if} statements were modeled using \pitdc,
which replaced the boolean condition with true, which removed the effect
of conditional. The \textit{switch} statement deletion was modeled using
\pitrs which replaced the first 100 labels with default label,
resulting in the switch element getting deleted. (Due to the constraints
of the architecture of PIT only the first 100 labels were replaced). The
\textit{try catch} was not necessary at bytecode level.  Finally, to get an
accurate estimation of the effect of true statement deletion, we grouped
the mutants by line, and considered a single virtual mutant (That is, the
total number of mutants is equal to the total number of lines). Further,
killing any mutant from a line resulted in marking the virtual mutant
for that line as killed. This gave us the mutation score for the virtual
statement deletion operator.  We note that the approximation of simple
statement deletion, especially when arithmetic operators are involved is not
complete. However, as seen in Figure~\ref{fig:op}, the number of mutants
produced by \pitm operator is very small. We also note that our
approximation does not account for the increase in ease of detection when
multiple mutations are combined together due to coupling effect. This also
means that not all the lines may be mutated since there may be no applicable
operators. However, given the constraints of a bytecode based mutation system,
we believe that our procedure is reasonable.

>>>>

\section{Methodology}
\label{sec:methodology}
Our selection of programs was driven by a few overriding concerns. Our primary
requirement was that, our results had to be as widely applicable as possible
for real-world programs. Secondly, we strived for a statistically significant
result, reducing the amount of variables present in the experiments. For this
reason we chose a random sample of Java projects from Github\cite{github}
that uses the popular maven\cite{maven} build system, that has been used
elsewhere\cite{gopinath2014code}. We utilize two flavors of
of test-suites. The original test-suites present in a limited number of
projects, and automatically generated test-suites using randoop.  Due
limited resources, we had to further reduce the programs and test-suites
to those that could finish the mutation run in 10 minutes. From this set
of test-suites, we sampled 110 suites each for both original and randoop
generated test-suits. This gave us 81 unique projects for both original
and randoop, and remaining projects having both flavors of test suites.

@@

For a detailed description of each mutants, please see~\cite{pitmut}.
PIT was also modified to provide random selection of mutants according
to various criteria. Finally, to remove random noise, each criteria
described has been run four times, and the results are averaged to produce
the final result.

\begin{table}
\small{
\begin{tabular}{ll}
\hline \hline \\
\pitin & Remove negative sign from numbers\\
\pitrv & Mutate return values\\
\pitm & Mutate arithmetic operators\\
\pitvmc & Remove void method calls\\
\pitnc & Negate conditional statements\\
\pitcb & Modify boundaries in logical conditions\\
\piti & Modify increment and decrement statements\\
\pitnvmc & Remove non-void method calls, returning default value \\
\pitcc & Replace constructor calls, returning null \\
\pitic & Replace inline constants with default value \\
\pitri & Remove increment and decrement statements \\
\pitemv & Replace member variable assignments with default value \\
\pites & Modify switch statements \\
\pitrs & Replace switch labels with default (thus removing them) \\
\pitrc & Replace boolean conditions with true\\
\pitnc & Replace boolean conditions with false\\
\hline
\end{tabular}
}
\caption{PIT Mutation Operators (We have used abbreviations instead of operator names.)}
\label{tbl:pitop}
\end{table}

One of the criteria we sampled was \textit{statement deletion}.
Since we use PIT, which is a bytecode based mutation engine, implementing direct equivalent of
source code statement deletion is very hard. We were able to approximate statement deletion
by combining different bytecode operators, which is explained in Subsection~\ref{subsec:stmtdel}.

\subsection{Sampling Criteria}

@@

Next, we explored in the other direction, by forcing the program elements
to return at least one element by using the \textit{ceil(sample)}. That is,
irrespective of the fraction being sampled, the simplest elements always
contributed at least one mutant to the total sample.

@@

\subsubsection{Statement Deletion:}
\label{subsec:stmtdel}
The basic statement deletion was modeled on the work by Deng et
al.~\cite{deng2013empirical}. The operations on single statements was modeled
using \pitvmc, \pitnvmc, \pitcc, \pitemv, \pitri
for simple statements, and using \pitrc for control structures,
which replaced the boolean condition with false, resulting in removal of
the block of statements. The operator for return values was modeled using
\pitrv, which is similar.  The operators for \textit{while,for},
and \textit{if} statements were modeled using \pitdc,
which replaced the boolean condition with true, which removed the effect
of conditional. The \textit{switch} statement deletion was modeled using
\pitrs which replaced the first 100 labels with default label,
resulting in the switch element getting deleted. (Due to the constraints
of the architecture of PIT only the first 100 labels were replaced). The
\textit{try catch} was not necessary at bytecode level.  Finally, to get an
accurate estimation of the effect of true statement deletion, we grouped
the mutants by line, and considered a single virtual mutant (That is, the
total number of mutants is equal to the total number of lines). Further,
killing any mutant from a line resulted in marking the virtual mutant
for that line as killed. This gave us the mutation score for the virtual
statement deletion operator.  We note that the approximation of simple
statement deletion, especially when arithmetic operators are involved is not
complete. However, as seen in Figure~\ref{fig:op}, the number of mutants
produced by \pitm operator is very small. We also note that our
approximation does not account for the increase in ease of detection when
multiple mutations are combined together due to coupling effect. This also
means that not all the lines may be mutated since there may be no applicable
operators. However, given the constraints of a bytecode based mutation system,
we believe that our procedure is reasonable. An interesting point is that
by this procedure statement deletion is no longer a simple mutation and qualifies
as a higher order mutation operator. In fact, the procedure we outlined can
be used in general to judge the effectiveness of a higher order operator from
the results of first order mutation operators that makeup the higher order
mutant in question.
